---
title: "GetCleanDataAssigment Codebook"
author: "Me"
date: '2018-11-28'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Data Set Information:

The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain.

The test and train data sets were merged into one data set. Only the mean and standard deviation were extracted, and the data was given tidy data principles:

1. Each variable forms a column.
2. Each observation forms a row.
3. Each type of observational unit forms a table.

A tidy data set was created with the average of each variable for each activity, for each subject.

Intertial signals from the test and train data set were not used because they were not relevant to the analysis.

```{r cars}
run_analysis <- function (x) {
  
  #Load in training data sets
  
  subjectrain <- read.table("subject_train.txt")
  trainset <- read.table("X_train.txt")
  trainlabel <- read.table("y_train.txt")
  
  #Load in test data sets
  
  subjectest <- read.table("subject_test.txt")
  testset <- read.table("X_test.txt")
  testlabel <- read.table("y_test.txt")
  
  #Load in activity labels and put in descriptive columns
  
  activitylabels <- read.table("activity_labels.txt")
  actname <- c("Key","ActivityLabel")
  colnames(activitylabels) <- actname
  
  #Adhere to tidy data by putting a descriptive name for column headers of subject ID data set and activity data set
  colnames(subjectrain)[1] <- "SubjectID"
  colnames(subjectest)[1] <- "SubjectID"
  colnames(trainlabel)[1] <- "Activity"
  colnames(testlabel)[1] <- "Activity"
  
  #Adhere to tidy data by putting a descriptive name for column header of combined data set
  
  features <- read.table("features.txt")
  featuresvec <- as.character(features[,2])
  colnames(testset) <- featuresvec
  colnames(trainset) <- featuresvec
  
  #Bind Activities and subject IDs to the test and training sets
  
  trainset <- data.frame(subjectrain, trainlabel, trainset)
  testset <- data.frame(subjectest, testlabel, testset)
  
  #Bind test and training data
  combinedset <- rbind(testset,trainset)
  
  #Select only columns with "mean" and "std"
  patterns <- c("mean","std")
  parsedcombined <- combinedset[,c(1,2,grep(paste(patterns, collapse="|"),names(combinedset)))]
  
  #Replace activity label key with activity labels in combined dataset
  
  parsedcombined$Activity <- activitylabels$ActivityLabel[parsedcombined$Activity]
  
  #Make column names more readable; remove abbreviations
  
  colnames(parsedcombined) <- sub("tBody", "timeBody", colnames(parsedcombined))
  colnames(parsedcombined) <- sub("fBody", "frequencyBody", colnames(parsedcombined))
  colnames(parsedcombined) <- sub("Mag", "Magnitude", colnames(parsedcombined))
  colnames(parsedcombined) <- sub("Acc", "Accelerometer", colnames(parsedcombined))
  
  #Arrange by activity
  
  parsedcombined <- arrange(parsedcombined, Activity)
  
  #Group by Subject ID and Activity, then get the mean of each activity for each subject
  
  tidyset <- parsedcombined %>% group_by(SubjectID, Activity)
  tidyset2 <- tidyset %>% summarise_at(3:81,mean)
  
}
```


